---
title: "Meta-Analysis in R"
author: "Pedro Henrique Brant"
date: "`r Sys.Date()`"
output: html_document
---

We're going to practice doing a meta-analysis in R. First, let's load the necessary packages.

```{r, results = FALSE, warning = FALSE, message=FALSE, R.}
## checking if packages are installed, if they aren't, installing them 
## loading them after

if(!require(tidyverse)){
        install.packages("tidyverse")
}
library(tidyverse)

if(!require(meta)){
        install.packages("meta")
}
library(meta)

if(!require(readxl)){
        install.packages("readxl")
}
library(readxl)

if(!require(googlesheets4)){
        install.packages("googlesheets4")
}
library(googlesheets4)
```

We're going to read the extracted data from the studies directly from googlesheets. Two outcomes will have their results summarized in this analysis, one dichotomous and one continuous.

```{r, results = FALSE, message = FALSE, warning = FALSE}

## checking if the data has been read already, if it hasn't reading it
## from googlesheets directly

if(!exists("df", envir = .GlobalEnv, inherits = FALSE)){
        df <- read_sheet("https://docs.google.com/spreadsheets/d/12mRnNd5HrbpE54d0fL_AtFnlum86g8X5k8MoZRtc92w/")
}

## reading the dataset for continuous outcome
if(!exists("df1", envir = .GlobalEnv, inherits = FALSE)){
        df1 <- read_sheet("https://docs.google.com/spreadsheets/d/1wjJVGk-KobXbmKuhFDC0TIBaRLgIEO9-pV0zDmqQQQ4/")
}
```

Let's take a look at the data:

```{r}

head(df)
head(df1)

```

Now, creating the variables for both meta-analyses, the binary and the continuous.

```{r}
## meta-analysis variable for the binary outcome, mantel haenszel method
## reporting Relative risk
test1 <- metabin(evtto, ntto, evcont, ncont, estudo, data = df, sm = "RR", method = "MH")

## creating a meta-analysis variable for the continous variable, reporting 
## Standard Mean Differences (Hedges's g is the default)

test2 <- metacont (n.e, mean.e, sd.e, n.c, mean.c, sd.c, study, data = df1,
                   sm = "SMD")

```

With these variables we can create forest plots. One for each analysis:

```{r, fig.width = 10,fig.height=3}

test1 %>% forest(random = TRUE, common = FALSE, label.c = "Placebo",
                 label.right = "Favours Placebo", label.left = "Favours TXA",
                 label.e = "TXA", sortvar = 1/w.random)

test2 %>% forest (sortvar = 1/w.random, common = FALSE, random = TRUE,
                  label.e = "TXA", label.c = "Placebo",
                  label.right = "Favours Placebo", label.left = "Favours TXA")

```

In both cases, TXA was better than placebo.